\chapter{Discussion}
\label{chap:discussion}

Designing useful visualizations for event sequences is challenging because of the persistent tension between representativeness and interactivity. As highlighted in~\autoref{chap:intro}, visualizations that aim to represent all events sacrifice interactivity, whereas systems designed for interactivity risk discarding or misrepresenting events. To ease the tension, in this dissertation, we claim \textbf{``useful visual solutions require treating visual design considerations and data management strategies as interconnected components, rather than separate concerns.''} This chapter provides discussions on how an interconnected visual design and data management strategies make useful visualizations for event sequences by easing the tension between representativeness and interactivity.


\section{Connecting Visual Design Considerations with \\ Data Management Strategies}

In this section, we discuss considerations for connecting visual design with data management strategies that emerged from our experiences and learned lessons. We synthesize the lessons learned across our studies and implementations, emphasizing how data abstraction, query optimization, and visualization strategies collectively influence interactivity and representativeness. We also discuss the limitations of a connected approach, acknowledging that while it can yield powerful, tailored solutions, it may reduce generalizability and require careful balancing of design goals.

\subsection{Relating Visual Data Abstraction with Data Formats}
\label{subsec:abstraction_format}

A fundamental requirement that emerged inevitably during our work is the need to consciously relate the \textit{visual data abstraction} chosen for a visualization design to the \textit{data format} used to store and transmit that data across system components. A data abstraction is the mapping of raw domain data to an abstract data type suitable for visualization~\cite{munzner2014vad}. For example, representing a set of pairwise relationships as a graph, or a sequence of timestamped occurrences as an ordered list of events. We define \textit{visual data abstraction} as the data abstraction used directly for visual encoding. A \textit{data format} is the concrete structural and syntactic encoding in which data is physically held and exchanged between system components, such as the visual rendering layer, a data query manager, and the underlying databases~\cite{kandel2011wrangling}. Visualization designers often treat visual data abstraction and data format as separate concerns~\cite{heer2006software, bigelow2020guidelines}. Visualization systems define their own internal data abstractions tied to visual marks, while databases and backends impose their own storage formats. 

When the data abstraction assumed by the visual design and the data format used by the data management components are poorly aligned, two categories of problems arise. First, data may be \textit{misrepresented}, values that are semantically continuous may be rendered as discrete, hierarchical relationships may be flattened and lost, or missing entries may be silently misinterpreted. Second, the system incurs unnecessary \textit{data conversion overhead}, every component boundary between a format-native store and an abstraction-native renderer requires a translation step, increasing latency, introducing potential for error, and complicating maintenance~\cite{battle2020structured, heer2023mosaic, yang2024optimizing, kandel2011wrangling}. These problems create a persistent risk of a mismatch between visual data abstraction and data formats. Recognizing and resolving this mismatch early in the design process is necessary to maintain interactivity and representativeness.

We notice this mismatch during the progression of our work across the Traveler and ESeMan. In Traveler, we utilized a \textit{linear data format} for visualization where each element of the linear structure maps into a visual mark: one to a darker box, zero to no box, and any value in between to a gray box. While storing the data, we utilized an \textit{interval tree}. Therefore, prior to rendering the data, it was required to convert the interval tree into a linear structure to maintain our existing design choices. Later, we adopted a \textit{summed area table} that gives a significant boost to the interactive performance, but that sacrifices accuracy. Our contribution with ESeMan adopted a \textit{hierarchical data structure} to control the accuracy, which also helps to tune the interactive performance by balancing the visual accuracy. Crucially, the hierarchy was designed to mirror the visual data abstraction, each level of the hierarchy corresponded to a distinct granularity of visual representation, so that data fetched at a given level arrived in a form that required minimal conversion before rendering. The result was a measurable reduction in data fetch time, a reduction in inter-component conversion cost, and the preservation of semantically important event attributes such as start time and end time that had been lost or approximated in earlier format choices.

To apply this lesson in practice, we recommend identifying \textit{visual data abstraction} and \textit{data format} and iteratively updating both to relate with each other as closely as possible. Stakeholder engagement and domain characterization~\cite{munzner2009nested, Sedlmair2012}, revealing latent data abstractions~\cite{bigelow2020guidelines}, data reconnaissance and task wrangling~\cite{crisan2019uncovering}, and data-first strategy~\cite{oppermann2020data} are utilized to identify \textit{visual data abstraction}. \textit{Data formats} are identified by profiling the raw data landscape like determining structural organization (flat tables, nested hierarchies, multi-table collections, temporal sequences, or graphs), enumerating attribute types, ranges, and edge cases, and documenting how data flows between system components such as databases, query managers, and rendering layers~\cite{kandel2011research, battle2020structured, yang2024optimizing}. Finally, to relate the visual abstraction to the format, the gap between the two is enumerated as syntactic, structural, or semantic transformation requirements, and the data architecture is adapted, through schema redesign, query API restructuring, or data structure selection, so that the format natively mirrors the visual abstraction and minimizes inter-component conversion~\cite{kandel2011research, walny2019data, kasica2020table, battle2020structured, yang2024optimizing, bigelow2020guidelines}. Where format and abstraction cannot be aligned at the storage level, dedicated middleware architectures can coordinate between the visualization specification and the backend data system~\cite{battle2020structured, heer2023mosaic, yang2024optimizing}; and where domain partners must be brought into the alignment process~\cite{williams2023data}, the guidelines of Bigelow et al.~\cite{bigelow2020guidelines} provide a framework for negotiating which latent abstractions to reveal and at what cost.

\subsection{Designing and Optimizing Queries Based on Visual Tasks}
\label{subsec:query_tasks}

A critical step in building event sequence visualization systems is to first identify the visual tasks that users need to perform. This principle resonates with Munzner's~\cite{munzner2009nested} widely adopted nested model of visualization design, in which characterizing domain tasks is the uppermost, and most foundational, level, with errors there inevitably cascading to every downstream design decision, including data abstraction and algorithm selection. Establishing the task space, therefore, provides a foundation for determining which queries are most important and where optimization efforts will have the greatest impact.

Recognizing that a single query can support multiple visual tasks creates opportunities to optimize both the data management and the visual interface. Recent works~\cite{heer2023mosaic, yang2024optimizing, vieth2023manivault} demonstrate that decoupling data processing from visual specification, where each view publishes its data needs as declarative queries that are then managed and automatically optimized by a central coordinator, yields scalable performance across multiple linked visualization components simultaneously.

Moreover, optimizing shared queries that serve multiple visual tasks provides compounded benefits, improving both interactivity and cross-view consistency. When a single query endpoint underlies multiple views, optimizing that query raises the performance of all views simultaneously, rather than requiring per-view tuning. It also ensures that all coordinated views operate from a shared data representation, which prevents subtle discrepancies that can otherwise arise when views issue separate, independently constructed queries~\cite{battle2020structured, vieth2023manivault}.

For example, in Traveler, both the Utilization view and the Gantt view are supported by the same query endpoint. Utilizing the same query endpoint allows the system to compute and fetch data while simultaneously serving multiple visual components. As a result, interactivity is improved because the same optimized query serves multiple needs, and consistency is maintained because all views are based on a shared data representation. More generally, designing queries around identified visual tasks enables visualization systems to avoid redundancy in data retrieval, reduce discrepancies between coordinated views, and focus optimization efforts on the queries that matter most to visualization users.

To apply this lesson in practice, after identifying the visual tasks using an established task taxonomy~\cite{brehmer2013multi, zinat2024multi}, group tasks by shared data requirements. For example, tasks drawing on the same entities, time ranges, or aggregation granularity are candidates for a shared query endpoint~\cite{heer2023mosaic, battle2020structured}. Next, design parameterized query endpoints scoped to each task cluster rather than to individual views~\cite{moritz2019falcon, heer2023mosaic}. Concentrate optimization efforts (indexing, pre-aggregation, and cachin, etc.) on the endpoints serving the most frequent or latency-sensitive tasks first, as misaligning optimization strategy with the actual task profile is a primary cause of poor interactive performance~\cite{battle2020structured, moritz2019falcon}. Finally, validate that all views sharing a query endpoint remain mutually consistent and meet latency thresholds across the full range of supported interactions.

\subsection{Visible Pixels Aware Data Abstraction}
\label{subsec:pixel_abstraction}

Another important design consideration is to anchor the visual data abstraction in the visible pixel space. A system can avoid unnecessary computation on data that cannot be distinguished at the current resolution. We adopt a \emph{pixel window} approach, where the visible space is divided into pixel windows. Data management strategies are then adopted to fetch only the data necessary for these pixel windows.

The adoptation of a pixel window strategy provides two key benefits. First, it bounds the amount of data that must be processed to the display resolution, ensuring that interaction cost depends on the number of pixels rather than the total number of events. Second, it allows multiple summarization techniques, such as aggregation or sampling, to be naturally framed in terms of pixel windows, making results directly interpretable within the visual interface. In practice, the pixel window strategy guided the design of both our query layer and supporting data structures, ensuring that interactivity is maintained even at large scales.

The principle of pixel-aware abstraction has been highlighted in prior visualization systems, such as imMens~\cite{liu2013immens}, Falcon~\cite{moritz2019falcon}, and Nanocubes~\cite{lins2013nanocubes}, all of which demonstrate that performance can be decoupled from raw data size by constraining queries to the visible resolution. Our work extends the pixel-aware abstraction principle to event sequence visualization, where queries aligned with pixel windows allow interactive navigation of timelines while preserving representativeness at the visual level. The pixel-based framing exemplifies the central thesis of the dissertation: useful visual solutions require connecting visual designs with data management strategies, here by binding queries and data structures directly to the constraints of the visual interface.


\subsection{Balancing Representativeness and Interactivity}
\label{subsec:balance}

A central challenge in event sequence visualization arises from the need to satisfy both representativeness and interactivity~\cite{guo2021survey, ulmer2023survey}. Representativeness requires that the visualization accurately reflects the underlying data, while interactivity demands that the system responds quickly, within 100ms~\cite{battle2019role}, to user interaction. Achieving both simultaneously is difficult because ensuring representativeness often involves processing large volumes of data, which can lead to latency that undermines interactivity~\cite{battle2019role, moritz2017trust}. Conversely, optimizing for interactivity may require techniques that compromise representativeness~\cite{battle2020structured}.

Achieving both representativeness and interactivity simultaneously remains a difficult challenge for event sequence visualization. While our systems demonstrate that progress is possible, in practice, designers often need to balance between these two goals depending on user needs. One promising strategy is to provide options that allow users to control the balance explicitly. In our visual design, users control the data accuracy with the zoom level by scrolling and panning. Data queries are therefore designed to fetch specific data based on the zoom level and panning window. Although we didn't implement full progressive refinement~\cite{stolper2014progressive}, our system is designed to render an initial view quickly with approximate or partial data, and then refine the visualization with more accurate data as computation and data transfer complete. Our design ensures that the user receives immediate feedback, even if initially approximate, while progressively improving the representation. Visualizing all data at once is rarely useful for exploratory analysis; instead, the goal should be to present data that is useful for exploration while maintaining interactive responsiveness~\cite{battle2020structured, jo2014livegantt}.

The challenges of balancing representativeness and interactivity have been observed and characterized across the visualization literature. Progressive visualization research characterizes the challenge as a tradeoff between result accuracy (representativeness) and computation speed (interactivity), showing that completely blocking strategies (maximal representativeness but zero interactivity) significantly hinder user exploration compared to instantaneous or progressive approaches~\cite{zgraggen2016progressive}. Scalability studies~\cite{richer2022scalability} model this challenge as competing effort functions that system designers must explicitly balance. Optimistic visualization system in~\cite{moritz2017trust} showed that users can extract valid insights from approximate and progressively refined results as long as approximation quality (representativeness) remains above a minimum threshold and system latency (interactivity) stays low enough to maintain engagement.

In practice, before building a system, designers should characterize the minimum representativeness required for the analytic task to be trustworthy, and the maximum latency the workflow can absorb before exploration stalls~\cite{liu2014latency,richer2022scalability, battle2020structured, ulmer2023survey}. Furthermore, designers must choose an appropriate tradeoff depending on task requirements and the context in which the visualization will be used~\cite{zgraggen2016progressive, guo2021survey, richer2022scalability}. Finally, the level of representativeness and interactivity should be made user-controllable, allowing users to adjust the balance based on their current needs and preferences~\cite{tao2019kyrix, chen2021pyramid, heer2023mosaic, ulmer2023survey, hografer2023tailorable}. By explicitly acknowledging and managing the balance between representativeness and interactivity, designers can create visualizations that are useful for the exploration of large event sequences.

\subsection{Limitations of a Connected Approach}

While connecting visual design considerations with data management strategies brings substantial value, a connected approach has several limitations. One limitation is that a connected design often leads to highly tailored solutions, systems that are deeply optimized for a specific type of visualization or data structure. The considerations discussed in this section make a visualization system powerful within its intended domain but reduce its generalizability across different visualization contexts or datasets.

Another difficulty lies in anticipating what to prioritize between representativeness and interactivity. In some cases, maintaining high visual representativeness can impose significant data management overhead, reducing interactivity. In others, data constraints limit the level of visual detail that can be feasibly rendered. Balancing between representativeness and interactivity in a scalable manner is inherently complex and context-dependent~\cite{richer2022scalability}.

These limitations highlight the necessity for careful consideration of the specific goals and use-cases of a visualization system~\cite{munzner2009nested, Sedlmair2012}, the nature of the data being represented~\cite{williams2023data}, and the needs of users to balance customization and generalizability~\cite{Sedlmair2012, bullock2022let, subramonyam2023we}.

\section{Visual Design Guidelines}

In this section, we propose visual design guidelines in this section for creating useful event sequence visualizations by connecting visual design considerations with data management strategies. Then, we discuss how these guidelines address the design challenges identified in prior event sequence visualization research.


\subsection{Use Connected Multi-Level Data Abstraction}
\label{subsec:connected_multi_level_abstraction}

A visualization system is composed of multiple components, each serving a distinct role in the visualization pipeline from raw data to visual rendering. Components can manage data, render the visual interface, or occupy intermediate roles that mediate between these two ends. Each component maintains a data abstraction level, which defines how data is represented and structured within that component. Data abstraction levels vary across components, a data management component may operate on raw records, while a visual interface component works with visual interface-ready representations such as coordinates or colors. A connected multi-level data abstraction, our first guideline, ensures each data transformation~\cite{kandel2011research} between abstraction levels occurs at most once in the entire pipeline. Furthermore, data abstraction levels should be defined such that data can be directly mapped between adjacent components without requiring additional conversion.

Prior work has explored how to connect abstraction levels in visualization pipelines. The visual entity budget~\cite{elmqvist2009hierarchical} formalizes a connection by bounding the number of renderable entities at the current display resolution. ImMens~\cite{liu2013immens} unified abstraction levels through data tiles that map directly to binned visual summaries at fixed resolutions. Kyrix~\cite{tao2019kyrix} makes the connection declarative by attaching an explicit data transform specification to each zoom-level canvas. VegaFusion~\cite{kruchten2022vegafusion} partitions Vega transformation graphs between client and server by statically analyzing each transform's data requirements, ensuring that only the data granularity demanded by the current visual specification is computed and transferred. Mosaic~\cite{heer2023mosaic} introduces a coordinator that receives declarative queries from each visualization client and automatically optimizes data fetching against a backing database. For event sequence data specifically, Gotz et al.~\cite{gotz2019visual} show that a predefined event-type hierarchy can drive dynamic aggregation at runtime, allowing the data layer to recalibrate its output in direct response to the user's current exploration depth. Despite their contributions, none of these approaches enforces a strict limit on the number of transformations permitted between components.

A connected data abstraction has three concrete benefits. First, it eliminates redundant transformations, improving efficiency. Second, it prevents the common pitfall of treating data and visual abstraction as separate concerns~\cite{battle2020structured, zinat2023comparative}, ensuring that what appears on screen faithfully reflects how data is processed. Third, it preserves representativeness: users can clearly see whether they are viewing a summary or complete data, avoiding misinterpretation.
 
% In ESeMan, for example, the hierarchical data structure is explicitly aligned with visual zoom levels. Each hierarchy level corresponds to the resolution and granularity visible at that scale, allowing the system to fetch data ready for direct rendering without further transformation.
 
In practice, enforcing a connected multi-level abstraction requires careful upfront alignment of abstraction levels across all components, which is non-trivial. Visualization designers are therefore encouraged to follow established visualization process and validation frameworks during system design. Traditional models~\cite{munzner2009nested, meyer2015nested} prioritized a ``downstream'' approach to process and validation where data abstraction is fixed before visual encoding and algorithmic decisions are made. For large event sequences, data cannot be effectively abstracted before the visual and implementation constraints are fully realized~\cite{zinat2023comparative}. The ``downstream'' rigidity disconnects data abstraction from visual and implementation realities, risking data misrepresentation and reduced interactivity. Visualization designers are therefore encouraged to define a connection iteratively in the \emph{discover}, \emph{design}, and \emph{implement} stages of the nine-stage framework~\cite{Sedlmair2012}, and in the \emph{understand} and \emph{ideate} activities of the design activity framework~\cite{mckenna2014design}.

Connected multi-level data abstraction carries an inherent risk of being defined broadly to accommodate every possible query and visual task, which can lead to over-generalization. In the extreme case, a fully general abstraction degenerates into storing raw records at every level, negating the efficiency gains that the guideline was intended to provide. The second guideline is presented to mitigate over-generalization.

%-----------------------------


% Large event datasets cannot be meaningfully explored without scalable mechanisms for shifting between overview and detail. Visual interfaces should therefore provide intermediate visual data abstraction, presenting data in various contexts and supporting user needs. The components that fetch data should be aware of these visual transitions and adjust the data abstraction accordingly. A common pitfall in visualization systems is to treat data abstraction for visual presentation and data abstraction for managing the data as separate concerns~\cite{battle2020structured, zinat2023comparative}.

% To address the pitfall, a connected multi-level abstraction ensures consistency of the levels of detail in data across different components, what appears on screen and how data is processed. The connection between data abstractions also enhances efficiency by allowing the system to fetch only the data necessary for the current visual abstraction without any data transformations. Additionally, a connected data abstraction helps designers maintain transparency by clearly indicating how each visual abstraction is derived, preserving representativeness and preventing users from mistaking summaries for complete data. In ESeMan, for example, the hierarchical data structure is explicitly aligned with visual zoom levels. Each level in the hierarchy corresponds to the resolution and granularity visible at that scale. This alignment allows the system to fetch data that is directly suitable for rendering at each zoom level, eliminating the need for additional transformations and ensuring that users see a consistent representation of the data as they navigate through different levels of detail.

% In practice, achieving a connection between data abstraction requires maintaining shared parameters, like the current zoom level, viewport extent, or pixel resolution. These parameters can propagate upstream through the abstraction from the data management level to the visual levels. Authors in~\cite{elmqvist2009hierarchical} formalize the connection using the \emph{visual entity budget}, the maximum number of visual entities renderable at the current display resolution. Additionally, imMens~\cite{liu2013immens} established connection using data tiles that correspond directly to binned visual summaries at fixed resolutions. Kyrix~\cite{tao2019kyrix} makes the connection declarative by attaching an explicit data transform specification to each zoom-level canvas. VegaFusion~\cite{kruchten2022vegafusion} partitions Vega transformation graphs between client and server by statically analyzing each transform's data requirements, ensuring that only the data granularity demanded by the current visual specification is computed and transferred. Mosaic~\cite{heer2023mosaic} generalizes this further through a coordinator that receives declarative queries published by each visualization client and automatically optimizes data fetching against a backing database. For event sequence data specifically, Gotz et al.~\cite{gotz2019visual} demonstrate that a predefined event-type hierarchy can drive dynamic aggregation at runtime, allowing the data layer to continuously recalibrate its output in direct response to the user's current visual exploration depth.

% While these solutions reduce unnecessary data conversion and tighten the coupling between visual and data abstraction levels, none of them imposes a strict restriction on the number of transformations permitted between the components. Enforcing this in practice, however, is non-trivial, as it requires careful upfront alignment of data abstraction levels across all components. Visualization designers are 
% therefore encouraged to follow established visualization process and validation frameworks during system design.

% The visualization design theory has historically prioritized a hierarchical structured approach to process and validation. Visualization design models~\cite{munzner2009nested, meyer2015nested} established a ``downstream'' decision process where data abstraction, in the 2nd nested block of their proposed model, serves as the immutable prerequisite for visual encoding and algorithmic implementation in the 3rd and 4th nested block, respectively. For large event sequences, data cannot be effectively abstracted before the visual and implementation constraints are fully realized~\cite{zinat2023comparative}. The ``downstream'' rigidity disconnects data abstraction from visual and implementation realities, risking data misrepresentation and reduced interactivity. Our design guideline advocates for a connected multi-level data abstraction that needs to be defined using an iterative process. The nine-stage~\cite{Sedlmair2012} and the design activity~\cite{mckenna2014design} framework proposed iterative and overlapping activities for visualization design process. This guideline is applicable in the \emph{discover}, \emph{design}, and \emph{implement} stages of the Nine-stage framework. In the design activity framework, this guideline should be followed during \emph{ideate} and \emph{make} activities.

\subsection{Prioritize Visual Tasks Before Optimizing Data Management}
\label{subsec:user_tasks}

% Prior visualization research~\cite{munzner2009nested, brehmer2013multi, munzner2014vad} has emphasized that design should start with user tasks rather than data structures or formats. Building on this foundation, this dissertation shows that identifying tasks not only guides visual design but also informs the choice of data management strategies. By optimizing queries that directly map to frequent tasks, our systems achieve interactivity while maintaining representativeness.

% \subsection{Addressing Design Challenges Using The Proposed Guidelines}

% Event sequence visualization research has identified several persistent challenges related to scalability, interactivity, and maintaining visual accuracy. Surveys conducted in~\cite{yeshchenko2024survey, guo2021survey} revealed that most existing systems focus primarily on visual representations while treating data management and query mechanisms as secondary concerns, resulting in limited scalability when handling large collections of event sequences. Additionally, preserving attributes of events while maintaining temporal context is a common challenge, as many visualizations either sacrifice detail for clarity or become cluttered when attempting to show all attributes~\cite{van2023survey, magallanes2021sequen}. The guideline G1 helps maintain temporal context for different attributes with connected abstractions in different levels across visual and data management components.

% Furthermore, the lack of explicit connections between user tasks and underlying data management strategies has been highlighted in~\cite{battle2020structured, battle2019role, zinat2023comparative}, hampering interactivity. The guideline G2 addresses the interactivity issue by emphasizing the importance of identifying visual tasks first and then optimizing data management strategies to support those tasks, ensuring that the system remains both interactive and representative of the underlying data. Finally, while some systems introduce multi-level abstractions to manage scale, they often do so without ensuring consistency across levels, which can lead to user confusion and misinterpretation of data~\cite{krause2023visually, magallanes2021sequen}. G1 facilitates achieving both consistency and efficiency with the connected abstractions. By explicitly connecting visual design considerations with data management strategies, the guidelines (G1 and G2) overcome the limitations identified in prior event sequence visualization research, ultimately leading to more useful visualization systems.

The second guideline is to prioritize the visual tasks users must perform. The prioritization of tasks in visualization design has long been recognized.
Shneiderman's information-seeking mantra (overview first, zoom and filter, details on demand)~\cite{shneiderman2003eyes} implicitly encodes three distinct task types, each demanding a different data type from the system. Munzner's nested model~\cite{munzner2009nested} places task and data abstraction in its second block, above encoding and algorithm decisions, establishing tasks as the anchor from which all downstream design choices derive. Brehmer and Munzner~\cite{brehmer2013multi} later formalized a multi-level typology of abstract visualization tasks, providing a vocabulary that designers can use to systematically enumerate user goals before committing to any data representation. Aigner et al.~\cite{aigner2011visualization} and Schulz et al.~\cite{schulz2013design} demonstrate that the diversity of time-oriented tasks (comparison, pattern detection, anomaly identification) drives fundamentally different requirements for how temporal records must be indexed and aggregated.

The consequence of ignoring task priorities is well documented in the database-driven visualization literature. The lack of explicit connections between user tasks and underlying data management strategies has been highlighted in~\cite{battle2020structured, battle2019role, zinat2023comparative}, hampering interactivity.
Systems that pre-cache or pre-aggregate data without grounding those decisions in actual query patterns routinely waste resources on rarely executed paths while leaving the most frequent interactions under-supported. Moritz et al.~\cite{moritz2019falcon} profile interaction logs to identify which crossfilter queries are most latency-sensitive and then selectively materialize only those index structures, leaving less critical queries to be answered on-the-fly. Similarly, ImMens~\cite{liu2013immens} restricts its data tile vocabulary to the binned aggregations demanded by its target tasks (brushing and linking across large multidimensional datasets) rather than attempting to serve arbitrary queries.

% For event sequence visualization, common user tasks include exploring event distributions over time, comparing utilization across resources, and examining dependencies among events. The Gantt Chart Task Taxonomy (\autoref{chap:gantt-taxonomy}) formalized many of these tasks, providing a structured framework for reasoning about task demands. Traveler (\autoref{chap:traveler}) exemplifies the guideline G2 by aligning multiple coordinated views with distinct task categories. Similarly, ESeMan tailored the query design to range and conditional range queries, as these directly map to the most frequent and useful Gantt tasks.

In practice, designers should identify user goals and the visual tasks that serve those goals. Subsequently, data management strategies should be identified to support the visual tasks. Starting from visual tasks provides a clear rationale for which queries must be efficient, what abstractions are needed, and where optimization efforts should be focused. Without the rationale, data management decisions risk over-optimizing for tasks that are rarely performed or under-supporting tasks that are critical to the analysis workflow. This guideline should be applied using an iterative design process, specifically, during the \emph{discover} and \emph{design} stages of the nine-stage framework~\cite{Sedlmair2012} and using the \emph{understand} and \emph{ideate} activities of the design activity framework~\cite{mckenna2014design}.

Additionally, when applying the second guideline, designers should confirm that each data abstraction is anchored to a concrete set of visual tasks, ensuring that the abstraction identified by the first guideline is also a \emph{minimal} one, retaining only the data the visual tasks actually demand.
